Analysis of Weight Lifting Exercise Dataset
========================================================

## Executive Summary
This analysis is conducted based on the Weight Lifting Exercise Dataset provided at http://groupware.les.inf.puc-rio.br/har. The goal of this project is to  use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants and then predict the manner in which they did the exercise.

## Preprocessing
First, let's load the data.
```{r}
training <- read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")

```

Examine the structure of the dataset:
```{r}
str(training)
```
Together with other functions such as dim(), head(), summary() ... I noticed there are quite a lot of empty strings and NAs in the training dataset, so I decided to remove all those columns.
We also should remove those columns that are not related to the classification (e.g. timestamp)

```{r}
training[training == '' | training == 'NA'] <- NA
trainingClean <- training[,complete.cases(t(training))]
trainingFinal <- trainingClean[,-(1:7)]
```

Now that we've cleaned the training dataset, we need to perform the same process on the testing dataset:
```{r}
testingClean <- testing[,names(testing) %in% names(trainingFinal)]
```

Before building our model, let's split the data into training and validation sets.
```{r}
library(caret)
inTrain <- createDataPartition(y=trainingFinal$classe, p=0.75, list=FALSE)
trainingSet <- trainingFinal[inTrain,]
validationSet <- trainingFinal[-inTrain,]
```


## Model Building
Initially, I tried to build a decision tree model. However, this accuracy is not great. As performance is not a critical requirement here, I decided to use random forests which can provide better accuracy.
```{r}
modelFit <- train(classe ~ ., method="rf", data=trainingSet, ntree=10, tuneLength=1)

```


## Cross Validation and Out of Sample Error
Now let's apply the model to validation dataset to understand the out of sample error of our model.
```{r}
confMatrix <- confusionMatrix(validationSet$classe, predict(modelFit, newdata=validationSet))
confMatrix
```
We can see the out of sample error on the validation set is rather low (<2%). Therefore, our model is relatively good.


## Prediction
Lets now predict using our random forests model.
```{r}
answers <- predict(modelFit, newdata=testingClean)
answers
```

